---
title: "Ubi_Kayu_mfg"
author: "Aulia Dini Rafsanjani"
date: "`r Sys.Date()`"
output: html_document
---

Dokumentasi ini dibuat sebagai bagian dari penyusunan neraca komoditas pada industri pati ubi kayu. 

Data mentah diterima dalam bentuk excel dengan nama mfg file. 

KBLI industri pati ubi kayu adalah 10621 dan 10613 

Series yang disajikan adalah tahun 2017-2022

Grafik yang ditampilkan yaitu :

1. Jumlah perusahaan 

2. Jumlah tenaga kerja

3. Total output. 

Output didekati dari nilai produksi. 

Pada penghitungan menggunakan file mfg, terdapat asumsi bahwa nilai produksi tidak hanya mencakup tepung tapioka saja tapi bisa saja ada nilai produksi dari hasil sampingan selain tepung tapioka. 

4. Sebaran jumlah perusahaan 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Library

```{r}
# Library 
library(readxl)
library(dplyr)
library(ggplot2)
library(sf)
library(stringr)
library(leaflet)
```

### Baca data

```{r}
# read dataset
stpim22 <- read_excel("/Users/auliadinirafsanjani/Documents/Industri/01 Agro Summary/Data STPIM/mfg22h.xlsx")
stpim21 <- read_excel("/Users/auliadinirafsanjani/Documents/Industri/01 Agro Summary/Data STPIM/mfg21e.xlsx", sheet = "Data")
stpim20 <- read_excel("/Users/auliadinirafsanjani/Documents/Industri/01 Agro Summary/Data STPIM/mfg20f.xlsx")
stpim19 <- read_excel("/Users/auliadinirafsanjani/Documents/Industri/01 Agro Summary/Data STPIM/mfg19E.xlsx")
stpim18 <- read_excel("/Users/auliadinirafsanjani/Documents/Industri/01 Agro Summary/Data STPIM/mfg18d.xlsx")
stpim17 <- read_excel("/Users/auliadinirafsanjani/Documents/Industri/01 Agro Summary/Data STPIM/mfg17e.xlsx")
```

### Filter data

```{r}
sub22_1 <- stpim22 %>% filter(Disic522 == "10621" | Disic522 == "10613")
sub21_1 <- stpim21 %>% filter(Disic521 == "10621" | Disic521 == "10613")
sub20_1 <- stpim20 %>% filter(Disic520 == "10621" | Disic520 == "10613")
sub19_1 <- stpim19 %>% filter(Disic519 == "10621" | Disic519 == "10613")
sub18_1 <- stpim18 %>% filter(Disic518 == "10621" | Disic518 == "10613")
sub17_1 <- stpim17 %>% filter(Disic517 == "10621" | Disic517 == "10613")
```

### Jumlah Perusahaan

```{r}
library(dplyr)
library(ggplot2)

# Create a list of datasets with corresponding years
data_list <- list(
  "2017" = sub17_1,
  "2018" = sub18_1,
  "2019" = sub19_1,
  "2020" = sub20_1,
  "2021" = sub21_1,
  "2022" = sub22_1
)

# Count number of rows (total establishments) for each year
row_counts <- lapply(names(data_list), function(year) {
  data <- data_list[[year]]
  tibble(
    Year = as.integer(year),
    Total_Establishments = nrow(data)
  )
}) %>%
  bind_rows()

# Print the summary table
print(row_counts)

# Create bar chart with dark orange bars, labels, and y-axis limit = 200
plot1 <- ggplot(row_counts, aes(x = factor(Year), y = Total_Establishments)) +
  geom_col(fill = "darkorange") +
  geom_text(aes(label = Total_Establishments),
            vjust = -0.5, size = 4, color = "black") +
  scale_y_continuous(limits = c(0, 200)) +  # changed from 180 to 200
  labs(
    title = "Jumlah Perusahaan Industri Pati Ubi Kayu tahun 2017–2022",
    x = "Tahun",
    y = "Jumlah Perusahaan"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

# print
print(plot1)

# Save the plot as JPEG
ggsave(filename = "plot1.jpeg", 
       plot = plot1, 
       width = 10, height = 6, units = "in", 
       dpi = 300)
```

### Tenaga Kerja

```{r}
# tk 2022
sub22_1$tk22 <- sub22_1$Lprnop22 + sub22_1$Lprnon22 + sub22_1$Lnpnop22 + sub22_1$Lnpnon22

# total labor 21
sub21_1$tk21 <- sub21_1$Lprnol21 + sub21_1$Lprnof21 + sub21_1$Lnpnol21 + sub21_1$Lnpnof21

# total labor 20
sub20_1$tk20 <- sub20_1$Lprnol20 + sub20_1$Lprnof20 + sub20_1$Lnpnol20 + sub20_1$Lnpnof20

# total labor 19
sub19_1$tk19 <- sub19_1$Lprnol19 + sub19_1$Lprnof19 + sub19_1$Lnpnol19 + sub19_1$Lnpnof19 

# total labor 18
sub18_1$tk18 <- sub18_1$Lprnol18 + sub18_1$Lprnof18 + sub18_1$Lnpnol18 + sub18_1$Lnpnof18 

# total labor 17
sub17_1$tk17 <- sub17_1$Ltmnou17 + sub17_1$Ltwnou17
```

```{r}
library(dplyr)
library(ggplot2)
library(scales)  # for label formatting

# Create a named list of datasets and their corresponding tk variable names
data_list <- list(
  "2017" = sub17_1$tk17,
  "2018" = sub18_1$tk18,
  "2019" = sub19_1$tk19,
  "2020" = sub20_1$tk20,
  "2021" = sub21_1$tk21,
  "2022" = sub22_1$tk22
)

# Compute total TK (sum) per year
tk_summary <- lapply(names(data_list), function(year) {
  tibble(
    Year = as.integer(year),
    Total_TK = sum(data_list[[year]], na.rm = TRUE)
  )
}) %>%
  bind_rows()

# Print summary table
print(tk_summary)

# Create bar chart with dark brown bars and thousand separators
plot2 <- ggplot(tk_summary, aes(x = factor(Year), y = Total_TK)) +
  geom_col(fill = "saddlebrown") +
  geom_text(aes(label = format(Total_TK, big.mark = ".", decimal.mark = ",")),
            vjust = -0.5, size = 4, color = "black") +
  scale_y_continuous(
    limits = c(0, 25000),
    labels = label_number(big.mark = ".", decimal.mark = ",")
  ) +
  labs(
    title = "Total Tenaga Kerja pada Industri Pati Ubi Kayu tahun 2017–2022",
    x = "Tahun",
    y = "Total Tenaga Kerja"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

plot2

# Save the plot as JPEG
ggsave(filename = "plot2.jpeg", 
       plot = plot2, 
       width = 10, height = 6, units = "in", 
       dpi = 300)
```

### Sebaran by provinsi tahun 2022

```{r}
# create map
# open the map file
geojson_data <- st_read("/Users/auliadinirafsanjani/Documents/Industri/01 Agro Summary/indonesia-prov.geojson")
```

```{r}
library(dplyr)
library(ggplot2)
library(sf)
library(viridis)

# 1️⃣ Compute total establishments by province
prov_join <- sub22_1 %>%
  group_by(Dprovi22) %>%
  summarise(count_22 = n()) %>%
  ungroup() %>%
  rename(kode = Dprovi22)  # make sure it matches the geojson 'kode' column

prov_join$kode <- as.integer(prov_join$kode)

# 2️⃣ Join with geojson map data
merged_data <- geojson_data %>%
  left_join(prov_join, by = "kode")

# 3️⃣ Create choropleth map
gbr1 <- ggplot(data = merged_data) +
  geom_sf(aes(fill = count_22), color = "white") +  # color = "white" for borders
  scale_fill_viridis_c(option = "plasma", na.value = "grey90") +  # nice color scale
  theme_minimal() +
  labs(
    title = "Peta Sebaran Industri Pati Ubi Kayu tahun 2022",
    fill = "Jumlah",
    caption = "Sumber: STPIM2023"
  ) +
  theme(
    plot.caption = element_text(hjust = 0, size = 10, face = "italic"),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )
gbr1

# Save the plot as JPEG
ggsave(filename = "gbr1.jpeg", 
       plot = gbr1, 
       width = 10, height = 6, units = "in", 
       dpi = 300)
```

### Total Output

```{r}
library(dplyr)
library(ggplot2)

# Create a named list of datasets with their corresponding output variable names
data_list <- list(
  "2017" = list(data = sub17_1, var = "Yprvcu17"),
  "2018" = list(data = sub18_1, var = "Yprvcu18"),
  "2019" = list(data = sub19_1, var = "Yprvcu19"),
  "2020" = list(data = sub20_1, var = "Yprvcu20"),
  "2021" = list(data = sub21_1, var = "Yprvcu21"),
  "2022" = list(data = sub22_1, var = "Yprvcu22")
)

# Compute sum for each year and convert to millions
output_summary <- lapply(names(data_list), function(year) {
  data <- data_list[[year]]$data
  var_name <- data_list[[year]]$var
  
  tibble(
    Year = as.integer(year),
    Total_output = sum(data[[var_name]], na.rm = TRUE) / 1e6  # divide by 1,000,000
  )
}) %>%
  bind_rows()

# Print summary
print(output_summary)

# Create line chart in dark grey with y-axis limit 10,000
plot3 <- ggplot(output_summary, aes(x = Year, y = Total_output)) +
  geom_line(color = "darkgrey", size = 1.2) +
  geom_point(color = "darkgrey", size = 3) +
  geom_text(aes(label = format(Total_output, big.mark = ".", decimal.mark = ",", nsmall = 2)),
            vjust = -0.5, size = 4) +
  labs(
    title = "Total Output per Tahun Industri Pati Ubi Kayu tahun 2017–2022",
    x = "Tahun",
    y = "Total output (Juta Rp)"
  ) +
  scale_x_continuous(breaks = 2017:2022) +
  scale_y_continuous(limits = c(0, 10000)) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

# Save the plot as JPEG
ggsave(filename = "plot3.jpeg", 
       plot = plot3, 
       width = 10, height = 6, units = "in", 
       dpi = 300)
```

```{r}
library(dplyr)
library(ggplot2)

# Create a named list of datasets with their corresponding output variable names
data_list <- list(
  "2017" = list(data = sub17_1, var = "Yprvcu17"),
  "2018" = list(data = sub18_1, var = "Yprvcu18"),
  "2019" = list(data = sub19_1, var = "Yprvcu19"),
  "2020" = list(data = sub20_1, var = "Yprvcu20"),
  "2021" = list(data = sub21_1, var = "Yprvcu21"),
  "2022" = list(data = sub22_1, var = "Yprvcu22")
)

# Compute sum for each year and convert to trillions (Rp)
output_summary <- lapply(names(data_list), function(year) {
  data <- data_list[[year]]$data
  var_name <- data_list[[year]]$var
  
  tibble(
    Year = as.integer(year),
    Total_output = sum(data[[var_name]], na.rm = TRUE) / 1e9
  )
}) %>%
  bind_rows()

# Function to format numbers with point as thousands and comma as decimals
format_label <- function(x) {
  sapply(x, function(value) {
    parts <- strsplit(sprintf("%.2f", value), "\\.")[[1]]
    integer <- as.integer(parts[1])
    decimal <- parts[2]
    integer_formatted <- format(integer, big.mark = ".", scientific = FALSE)
    paste0(integer_formatted, ",", decimal)
  })
}

# Plot
plot3 <- ggplot(output_summary, aes(x = Year, y = Total_output)) +
  geom_line(color = "darkgrey", size = 1.2) +
  geom_point(color = "darkgrey", size = 3) +
  geom_text(aes(label = format_label(Total_output)), vjust = -0.5, size = 4) +
  labs(
    title = "Total Output per Tahun Industri Pati Ubi Kayu (2017–2022)",
    x = "Tahun",
    y = "Total output (Triliun Rp)"
  ) +
  scale_x_continuous(breaks = 2017:2022) +
  scale_y_continuous(
    limits = c(0, 15),
    labels = format_label
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.text = element_text(size = 12),
    axis.title = element_text(size = 13)
  )

# print result
print(plot3)

# Save the plot as JPEG
ggsave(filename = "plot3.jpeg", 
       plot = plot3, 
       width = 10, height = 6, units = "in", 
       dpi = 300)
```


